\textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\ {\footnotesize Bachelorarbeit}

### 2023-04-16 Paper/chapters/implementation-strategies.tex
ps tabulation paper

\begin{description}
    \item[XOR cancellation property in key schedule:]
        \begin{align*}
            K_A&\leftarrow K_L\oplus K_R \\
            K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
            K_A&\leftarrow K_A\oplus K_L \\
        \end{align*}

@@

    \item[Absorption of whitening keys:] Key additions $kw_1,kw_3$ can be
        absorped into other subkeys:

>>>>

\begin{description}
    \item[XOR cancellation property in key schedule:]
        \begin{align*}
            K_A&\leftarrow K_L\oplus K_R \\
            K_A&\leftarrow FE(FE(K_A, \Sigma_0=\mathrm{0xa09e667f3bcc908b}), \Sigma_1=\mathrm{0xb67ae8584caa73b2}) \\
            K_A&\leftarrow K_A\oplus K_L \\
        \end{align*}

@@

    \item[Absorption of whitening keys:] Whitening keys $kw_1,kw_3$ can be
        absorbed into other subkeys to save two XOR operations

    \item[Efficiently computing $P\circ S$:] This technique applies to 64-bit processors.
        By preparing tables

        \begin{alignat*}{9}
            & SP_0(y_{0(8)})=(&s_0(y_0),&& s_0(y_0),&& s_0(y_0),&& 0,&& s_0(y_0),&& 0,&& 0,&& s_0(y_0)&) \\
            & SP_1(y_{1(8)})=(&0,&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& 0,&& 0&) \\
            & SP_2(y_{2(8)})=(&s_2(y_2),&& 0,&& s_2(y_2),&& s_2(y_2),&& 0,&& s_2(y_2),&& s_2(y_2),&& 0&) \\
            & SP_3(y_{3(8)})=(&s_3(y_3),&& s_3(y_3),&& 0,&& s_3(y_3),&& 0,&& 0,&& s_3(y_3),&& s_3(y_3)&) \\
            & SP_4(y_{4(8)})=(&0,&& s_2(y_4),&& s_2(y_4),&& s_2(y_4),&& 0,&& s_2(y_4),&& s_2(y_4),&& s_2(y_4)&) \\
            & SP_5(y_{5(8)})=(&s_3(y_5),&& 0,&& s_3(y_5),&& s_3(y_5),&& s_3(y_5),&& 0,&& s_3(y_5),&& s_3(y_5)&) \\
            & SP_6(y_{6(8)})=(&s_4(y_6),&& s_4(y_6),&& 0,&& s_4(y_6),&& s_4(y_6),&& s_4(y_6),&& 0,&& s_4(y_6)&) \\
            & SP_7(y_{7(8)})=(&s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0,&& s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0&)
        \end{alignat*}

        , we can compute $(P\circ S)(X_{(64)})$ with only 8 table lookups and 7
        XORs in the following way:

        \[
            (z_1', z_2', z_3', z_4', z_5', z_6', z_7', z_8')\leftarrow\bigoplus_{i=0}^7 SP_i(y_i)
        \]

### 2023-04-16 Paper/chapters/introduction.tex
ps tabulation paper

\begin{align*}
    K_A&\leftarrow K_L\oplus K_R \\
    K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
    K_A&\leftarrow K_A\oplus K_L \\
    K_A&\leftarrow FE(FE(K_A, \Sigma_2=0xc6ef372fe94f82be), \Sigma_3=0x54ff53a5f1d36f1c) \\
    K_B&\leftarrow FE(FE(K_A, \Sigma_4=0x10e527fade682d1d), \Sigma_5=0xb05688c2b3e6c1fd)
\end{align*}

@@

\begin{table}[h!]
    \centering
    \scriptsize
    \caption{Subkeys for 128-bit key}
    \begin{tabular}{lll|lll}
        \toprule
        Usage & Subkey & Value & Usage & Subkey & Value \\
        \midrule
        \multirow{2}{*}{Prewhitening} & $kw_{0(64)}$ & $(K_L{\lll_0})_{L(64)}$ & $F(\text{Round 9})$ & $k_{9(64)}$ & $(K_L{\lll_{60}})_{R(64)}$ \\
                                      & $kw_{1(64)}$ & $(K_L{\lll_0})_{R(64)}$ & $F(\text{Round 10})$ & $k_{10(64)}$ & $(K_A{\lll_{60}})_{L(64)}$ \\
        $F(\text{Round 0})$ & $k_{0(64)}$ & $(K_A{\lll_0})_{L(64)}$            & $F(\text{Round 11})$ & $k_{11(64)}$ & $(K_A{\lll_{60}})_{R(64)}$ \\
        $F(\text{Round 1})$ & $k_{1(64)}$ & $(K_A{\lll_0})_{R(64)}$            & $FL$        & $kl_{2(64)}$ & $(K_L{\lll_{77}})_{L(64)}$ \\
        $F(\text{Round 2})$ & $k_{2(64)}$ & $(K_L{\lll_{15}})_{L(64)}$         & $FL^{-1}$   & $kl_{3(64)}$ & $(K_L{\lll_{77}})_{R(64)}$ \\
        $F(\text{Round 3})$ & $k_{3(64)}$ & $(K_L{\lll_{15}})_{R(64)}$         & $F(\text{Round 12})$ & $k_{12(64)}$ & $(K_L{\lll_{94}})_{L(64)}$ \\
        $F(\text{Round 4})$ & $k_{4(64)}$ & $(K_A{\lll_{15}})_{L(64)}$         & $F(\text{Round 13})$ & $k_{13(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
        $F(\text{Round 5})$ & $k_{5(64)}$ & $(K_A{\lll_{15}})_{R(64)}$         & $F(\text{Round 14})$ & $k_{14(64)}$ & $(K_A{\lll_{94}})_{L(64)}$ \\
        $FL$        & $kl_{0(64)}$ & $(K_A{\lll_{30}})_{L(64)}$                & $F(\text{Round 15})$ & $k_{15(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
        $FL^{-1}$   & $kl_{1(64)}$ & $(K_A{\lll_{30}})_{R(64)}$                & $F(\text{Round 16})$ & $k_{16(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
        $F(\text{Round 6})$ & $k_{6(64)}$ & $(K_L{\lll_{45}})_{L(64)}$         & $F(\text{Round 17})$ & $k_{17(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
        $F(\text{Round 7})$ & $k_{7(64)}$ & $(K_L{\lll_{45}})_{R(64)}$         & \multirow{2}{*}{Postwhitening} & $kw_{2(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
        $F(\text{Round 8})$ & $k_{8(64)}$ & $(K_A{\lll_{45}})_{L(64)}$         &                                & $kw_{3(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
        \bottomrule
    \end{tabular}
\end{table}

Subkey generation for 192/256-bit keys are generated in a similar way.

>>>>

\begin{align*}
    K_A&\leftarrow K_L\oplus K_R \\
    K_A&\leftarrow FE(FE(K_A, \Sigma_0\mathrm{0xa09e667f3bcc908b}), \Sigma_1\mathrm{0xb67ae8584caa73b2}) \\
    K_A&\leftarrow K_A\oplus K_L \\
    K_A&\leftarrow FE(FE(K_A, \Sigma_2\mathrm{0xc6ef372fe94f82be}), \Sigma_3\mathrm{0x54ff53a5f1d36f1c}) \\
    K_B&\leftarrow FE(FE(K_A, \Sigma_4\mathrm{0x10e527fade682d1d}), \Sigma_5\mathrm{0xb05688c2b3e6c1fd})
\end{align*}

@@

\begin{table}[h!]
    \centering
    \scriptsize
    \caption{Subkey creation for 128-bit keys}
    \begin{tabular}{lll|lll}
        \toprule
        Usage & Subkey & Value & Usage & Subkey & Value \\
        \midrule
        \multirow{2}{*}{Prewhitening} & $kw_{0(64)}$ & $(K_L{\lll_0})_{L(64)}$ & $F(\text{Round 9})$ & $k_{9(64)}$ & $(K_L{\lll_{60}})_{R(64)}$ \\
                                      & $kw_{1(64)}$ & $(K_L{\lll_0})_{R(64)}$ & $F(\text{Round 10})$ & $k_{10(64)}$ & $(K_A{\lll_{60}})_{L(64)}$ \\
        $F(\text{Round 0})$ & $k_{0(64)}$ & $(K_A{\lll_0})_{L(64)}$            & $F(\text{Round 11})$ & $k_{11(64)}$ & $(K_A{\lll_{60}})_{R(64)}$ \\
        $F(\text{Round 1})$ & $k_{1(64)}$ & $(K_A{\lll_0})_{R(64)}$            & $FL$        & $kl_{2(64)}$ & $(K_L{\lll_{77}})_{L(64)}$ \\
        $F(\text{Round 2})$ & $k_{2(64)}$ & $(K_L{\lll_{15}})_{L(64)}$         & $FL^{-1}$   & $kl_{3(64)}$ & $(K_L{\lll_{77}})_{R(64)}$ \\
        $F(\text{Round 3})$ & $k_{3(64)}$ & $(K_L{\lll_{15}})_{R(64)}$         & $F(\text{Round 12})$ & $k_{12(64)}$ & $(K_L{\lll_{94}})_{L(64)}$ \\
        $F(\text{Round 4})$ & $k_{4(64)}$ & $(K_A{\lll_{15}})_{L(64)}$         & $F(\text{Round 13})$ & $k_{13(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
        $F(\text{Round 5})$ & $k_{5(64)}$ & $(K_A{\lll_{15}})_{R(64)}$         & $F(\text{Round 14})$ & $k_{14(64)}$ & $(K_A{\lll_{94}})_{L(64)}$ \\
        $FL$        & $kl_{0(64)}$ & $(K_A{\lll_{30}})_{L(64)}$                & $F(\text{Round 15})$ & $k_{15(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
        $FL^{-1}$   & $kl_{1(64)}$ & $(K_A{\lll_{30}})_{R(64)}$                & $F(\text{Round 16})$ & $k_{16(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
        $F(\text{Round 6})$ & $k_{6(64)}$ & $(K_L{\lll_{45}})_{L(64)}$         & $F(\text{Round 17})$ & $k_{17(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
        $F(\text{Round 7})$ & $k_{7(64)}$ & $(K_L{\lll_{45}})_{R(64)}$         & \multirow{2}{*}{Postwhitening} & $kw_{2(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
        $F(\text{Round 8})$ & $k_{8(64)}$ & $(K_A{\lll_{45}})_{L(64)}$         &                                & $kw_{3(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
        \bottomrule
    \end{tabular}
\end{table}

Subkeys for 192/256-bit keys are generated in a similar way.
